Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-09 article/article.Rmd
Commit 2024-06-09 22:17

According to this study, in the period 1997 -- 2005, the mean AOD at $340\,\text{nm}$
$0.403$ with a trend of $-3.8±0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta=0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$
in 2005).

Simulations with uvspec for the above Ångström coefficients, with WC of $`r
round(BR[year==1997, pw_avg_mm],1)`\,\text{mm}$ and $`r round(BR[year==2005,
pw_avg_mm],1)`\,\text{mm}$ for 1997 and 2005 respectively and for a SZA of $55^\circ$

By interpolating the resulting clear sky irradiance, we have a trend of
$+0.21\,\%$ per year for the clear sky irradiance.  For the period 2005 -- 2005 we
used the mean monthly values from AERONET, with a similar simulation scheme. We
computed the trend of clear sky radiation for each month of the year, which gave us a
mean clear sky radiation trend of $+0.14\,\%$ per year. We applied this long term
change to the clear sky irradiance, in order to create a more realistic
representation of the irradiance for the whole study period
(Figure\nobreakspace{}\ref{fig:CS-change}).

```{r CS-change, echo=FALSE, fig.cap="Simulated long term change of the clear sky irradiance relevant to its climatological values, due to changed of AOD."}
knitr::include_graphics("../images/P-CS-change-1.png")
```

@@

## CE Criteria investigation

Our main focus was to positively identify over irradiance events related to CEs.  For
this reason we had to select an appropriate set of atmospheric parameterization, in
Libradtran, and define an relevant threshold for the CE identification.  The
selection of the atmospheric parameterization will provide us with a reference, that
has a good agreement with the clear sky conditions. Additional, with the
implementation of the long term change of AOD, as described above, it will be able to align
well with most of the natural variation of the physical quantity.  The selection of
adequate threshold, will allow us to have a robust method to distinguish the CE
cases, which will compensate for the limited accuracy of the input data, and the
natural variability of the atmosphere.

First, we evaluated the performance of the modelled clear radiation, for each of
the atmospheric level of clearness we have created as reference, in relation to the
measured GHI data, and choose as a representative of the clear sky radiation, the
case where
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
These values represent a typical atmosphere in Thessaloniki, with low load of
aerosols and humidity, which are the main factor that attenuate the GHI, excluding
clouds.  To create the upper limit of CE identification, we used a value relative to
clear sky irradiance
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}).
\begin{equation}
\text{CE} : \text{E}_\text{i} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.

>>>>

According to this study, in the period 1997 -- 2005, the mean AOD at $340\,\text{nm}$
$0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta=0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).  Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and 
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively and for a SZA of $55^\circ$ reveal trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
The SZA of $55^\circ$ was chosen as representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean of clear sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

```{r CS-change, echo=FALSE, fig.cap="Simulated long term change in clear sky irradiance relative to the climatological values due to changes in AOD."}
knitr::include_graphics("../images/P-CS-change-1.pdf")
```

@@

## Criteria for the identification of CE events

In this study our main focus was to quantify over irradiance events related to CEs. A
key issue for achieving this goal is to  define a threshold for the CE
identification, representative of the clear-sky irradiance at the time of each GHI
measurement. This depends on the selection of the appropriate atmospheric
parameterization for the RTM simulations. The implementation of the long term change
of AOD, discussed in section\nobreakspace{}\ref{rtmcs}, allows capturing a large part
of the natural variability of clear-sky GHI. However, the short-term variability of
AOD cannot be taken adequately into account when using monthly values for the model
simulations. We tried different approaches to increase the robustness of the
methodology and compensate for the limited accuracy of the RTM input data and the
unpredictable natural variability of the atmosphere.

First, we evaluated the performance of the modelled $\text{GHI}_\text{ref}$ in
relation to the measured GHI for different levels of atmospheric clearness, by using
in the RTM the monthly climatological AOD and WC, less their respective standard
deviations.
the atmospheric level of clearness we have created as reference, in relation to the
measured GHI data, and choose as a representative of the clear sky radiation, the
case where
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
These values represent a typical atmosphere in Thessaloniki, with low load of
aerosols and humidity, which are the main factor that attenuate the GHI, excluding
clouds.  To create the upper limit of CE identification, we used a value relative to
clear sky irradiance
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}).
\begin{equation}
\text{CE} : \text{E}_\text{i} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.
